Novel lattice Boltzmann method for simulation of strongly shear thinning viscoelastic fluids

### 2024-02-21 NovelLBM.tex
clarify

We introduce the ratio between the zero-shear and the infinite-shear (i.e.\ solvent) viscosity as\RK{nope $\eta_p$ is not zero shear viscosity!}
\begin{align}
	R_\eta = \frac{\eta_\text{p}}{\eta_\text{s}}		\label{eq:R_eta}
\end{align}
which will turn out to be an important parameter to determine numerical stability of LB simulations.
\SG{ wo don't actually show that $R_ \eta  $ determines stability? }\RK{we do not. We could do a plot with $\alpha_s$ decreasing until the sim breaks if you want}
This ratio can similarly be defined for other models such as FENE-P or the inelastic, but shear-thinning Carreau-Yasuda model \cite{carreau1972, yasuda1979}.
For our alginate solution we obtain $R_\eta = \num{48.2\pm0.4e3}$, indicating that the solution viscosity at very low shear rates is more than four orders of magnitude larger than the solvent viscosity which represents a severe challenge to standard numerical algorithms.
That this characteristic is not unique to alginate, but is in fact a generic feature of cell carrier fluids used in bioprinting is shown in table~\ref{tab:bioinks} where rheological data from various literature sources have been compiled into a single format to determine $R_\eta$.
The \SI{0.49}{\percent} methyl cellulose solution used in RT-DC measurements, in contrast, features a much lower value of $R_\eta = \num{18.7\pm0.4}$.

@@

In the von Mises stress shown in figure \ref{fig:cellInShear_vM} the effect of the cell is clearly visible.
We observe four distinct regions of high stress radiating outwards from the cell body.
These are particularly pronounced along the longest axis of the deformed cell.
\SG{ connect to the ''viscoelastic wings'' paper \cite{Zhang.2020}}\RK{we have a sim with higher shear, this might be better here}

>>>>

We introduce the ratio between the zero-shear and the infinite-shear (i.e.\ solvent) viscosity as\RK{nope $\eta_p$ is not zero shear viscosity!}
\begin{align}
	R_\eta = \frac{\eta_\text{p}}{\eta_\text{s}}		\label{eq:R_eta}
\end{align}
which will turn out to be an important parameter to determine numerical stability of LB simulations.
\SG{ wo don't actually show that $R_ \eta  $ determines stability? }\RK{we do not. We could do a few simulations with $\alpha_s = 0$ and increasing $R_\eta$ until the sim breaks if you want}
This ratio can similarly be defined for other models such as FENE-P or the inelastic, but shear-thinning Carreau-Yasuda model \cite{carreau1972, yasuda1979}.
For our alginate solution we obtain $R_\eta = \num{48.2\pm0.4e3}$, indicating that the solution viscosity at very low shear rates is more than four orders of magnitude larger than the solvent viscosity which represents a severe challenge to standard numerical algorithms.
That this characteristic is not unique to alginate, but is in fact a generic feature of cell carrier fluids used in bioprinting is shown in table~\ref{tab:bioinks} where rheological data from various literature sources have been compiled into a single format to determine $R_\eta$.
The \SI{0.49}{\percent} methyl cellulose solution used in RT-DC measurements, in contrast, features a much lower value of $R_\eta = \num{18.7\pm0.4}$.

@@

In the von Mises stress shown in figure \ref{fig:cellInShear_vM} the effect of the cell is clearly visible.
We observe four distinct regions of high stress radiating outwards from the cell body.
These are particularly pronounced along the longest axis of the deformed cell.
\SG{ connect to the ''viscoelastic wings'' paper \cite{Zhang.2020}}\RK{Our "wings" look very different}
